Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-06 paper/disc20clustering.tex
wip

\section{Note on distributed algorithms}

\subsection{Distributed implementation of $k=3$ algorithm}

For a repartition, it is sufficient to find the cluster with the signature $1,1,1$. 
We call any such cluster \emph{X (todo name)}.
Idea of distributed implementation of this procedure:
\begin{enumerate}
	\item Assign ids to clusters and order them
	\item Recursively partition clusters into two groups
	\item Each group maintains the pointer to a subgroup that contains at least one X cluster; a pointer may be null (then no X there)
	\item This way, finding a cluster X requires $\log_2(\ell)$ communication
	\item After removing an X cluster, we need to fix the pointers. This is done by following the path from the removed X cluster to the root. We simply ask the sibling if it contains X.
	\item Possible improvement: we can have a fixed number of rounds for finding the X cluster by maintaining a direct pointer to an X cluster (not only to subgroup). However, the fix still requires the log number of rounds. But we improved the responsiveness.
\end{enumerate}

\subsection{Distributed implementation of $k=2$ algorithm}

Trivial? Counters are local. We migrate between the clusters that have just communicated.

\subsection{Distributed implementation of PPL}

\maciek{How to distribute PPL better?}

Our algorithms can be distributed
similarly to the approach in~\cite{sigmetrics19_partitioning}.

>>>>

\section{Distributed implementation.}
The algorithm PPL from Section~\ref{sec:ppl} can be distributed
similarly to the approach in~\cite{sigmetrics19_partitioning}.
The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration concerns only two clusters.

The algorithm for $k=3$ from Section~\ref{sec:k3} can be efficiently distributed.
There are two non-local operations:
\begin{enumerate}
	\item We broadcast the end of the phase to reset all counters.
	\item Merge of two  may require finding a third cluster that contains $3$ alone nodes (For the details, see the proof of Lemma~\ref{lem:1req})).
\end{enumerate}

If a component grows to size $4$ or fresh cluster is found, we conclude that no perfect partition exists, and we broadcast (\maciek{TODO: efficient broadcast cite}) the message to end the phase and reset the counters in all clusters.

Let's call the cluster containing $3$ alone nodes \emph{fresh}.
Now we show how to efficiently find a fresh cluster.
We organize the clusters into an arbitrary balanced binary tree (\maciek{TODO: how to do it efficiently? Does not matter much, as this is performed once at the beginning}), and broadcast the root.
Each cluster counts the number of fresh clusters in its subtree.
To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
Upon encountering a fresh cluster, we stop the traversal and decrease the counters on the followed path by $1$.

The operation (ref1) requires a single broadcast, and the operation (ref2) requires $O(\log(\ell))$ communication.
Each of these may occur on every request.

\maciek{Possible improvement: we can have a fixed number of rounds for finding the X cluster by maintaining a direct pointer to an X cluster (not only to subgroup). However, the fix still requires the log number of rounds. But we improved the responsiveness.}
